---
title: "Homework 4"
author: "Valentina Trevino"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(knitr)
library(tibble)
library(dplyr)
library(kableExtra)
library(mosaic)
```

### Problem 1 ###
```{r, warning=FALSE, echo=FALSE, message=FALSE}
setwd("/Users/valentinatrevino/Desktop/spring '24/SDS")
gas_data <- read.csv("gasprices.csv")
```

### Theory 1: Gas stations charge more if they lack direct competition in sight. ###

#### This theory claims that if there are competitors present the price of gas will most likely be higher than if there are no competitors. ####

```{r, warning=FALSE, echo=FALSE, message=FALSE}

mean(Price ~ Competitors, data= gas_data)
boot_comp = do(10000)*mean(Price ~ Competitors,
data=resample(gas_data))
diff_mean = (boot_comp$N - boot_comp$Y)
diff_mean_df <- data.frame(diff_mean)

ggplot(diff_mean_df) + 
  geom_histogram(aes(x=diff_mean), color= "deeppink", fill= "darkslategray") + labs(title = "Comparison of Mean Price of Gas Based on Competitors (Bootstrapped)", x= "Difference in Means") +
  scale_x_continuous(limits = c(-0.05, 0.085), breaks = seq(-0.05, 0.085, by = 0.01))
```

#### To investigate the potential correlation between gas prices and the presence of competitors at gas stations, I conducted a bootstrap analysis with ten thousand observations on two variables: the mean gas price at stations with competitors and those without. Subsequently, I constructed a histogram illustrating the difference in means between these two variables.

#### Upon examination of the histogram, it reveals a noticeable peak around 0.02 to 0.03. This indicates a slight increase in gas prices when competitors are present, albeit not a substantial one ####

### Theory 2: The richer the area, the higher the gas prices. ###

#### Claim: The second theory posits that the wealthier the area in which a gas station is situated, the higher the prices of gasoline sold there. ####

```{r, warning=FALSE, echo=FALSE, message=FALSE}
boot_gas <- replicate(10000, {
  sampled_data <- mosaic::resample(gas_data)
  cor(sampled_data$Income, sampled_data$Price)
})
boot_gas_df <- data.frame(boot_gas)

ggplot(boot_gas_df) + 
  geom_histogram(aes(x=boot_gas), color= "deeppink", fill= "darkslategray") + labs(title = "Distribution of Correlation Coefficients between Income and Gas Price (Bootstrapped)", x= "Correlation Coefficients")  + scale_x_continuous(limits = c(-0.1, 0.7), breaks = seq(-0.1, 0.7, by = 0.1)) 
```

#### Upon analyzing the histogram generated from the correlation coefficient between the "Income" and "Price" variables of the 10,000 sample datasets obtained through bootstrapping the gas_data, we observe a correlation of approximately 0.42 between these two variables.  ####

### Theory 3: Gas stations at stoplights charge more. ###

#### This theory claims that if a gas station is close to a stoplight, the price of the gas will most likely be higher than if it is not in a stoplight. ####

```{r, warning=FALSE, echo=FALSE, message=FALSE}

mean(Price ~ Stoplight, data= gas_data)
boot_stop = do(10000)*mean(Price ~ Stoplight,
data=resample(gas_data))
diff_mean2 = (boot_stop$Y - boot_stop$N)
diff_mean2_df <- data.frame(diff_mean2)

ggplot(diff_mean2_df) + 
  geom_histogram(aes(x=diff_mean2), color= "deeppink", fill= "darkslategray") + labs(title = "Comparison of Gas Price Means Based on Proximity to Stoplights (Bootstrapped)", x= "Difference in Means")
```

#### Upon examination of the histogram, it's evident that there isn't a clear correlation between gas prices and whether or not gas stations are located near a stoplight. This inference is drawn from the observation that the difference in mean prices between gas stations located at stoplights and those that are not is consistently close to zero. ####


### Theory 4: Gas stations with direct highway access charge more ###
#### This theory states that if a gas station has direct acces to a highway the price of the gas will be higher. ####

```{r, warning=FALSE, echo=FALSE, message=FALSE}
mean(Price ~ Highway, data= gas_data)
boot_high = do(10000)*mean(Price ~ Highway,
data=resample(gas_data))
diff_mean3 = (boot_high$Y - boot_high$N)
diff_mean3_df <- data.frame(diff_mean3)
ggplot(diff_mean3_df) + 
  geom_histogram(aes(x=diff_mean3), color= "deeppink", fill= "darkslategray") + labs(title =  "Comparison of Gas Price Means Based on Proximity to Highways (Bootstrapped)", x= "Difference in Means")
```

#### As observed in the histogram depicting the difference in mean prices between gas stations with highway access and those without, there is a peak difference around 0.05. This suggests that there is indeed a slight increase in price for gas stations with highway access. However, this increase is minimal, almost negligible. ####

### Theory 5: Shell charges more than all other non-Shell brands. ###
#### This theory states that Shell gas stations charge more than other gas stations like 7 - Eleven, Texaco, Gulf, among others. ####
```{r, warning=FALSE, echo=FALSE, message=FALSE}
mean(Price ~ Brand, data= gas_data)
boot_brand = do(10000)*mean(Price ~ Brand,
data=resample(gas_data))
boot_brand$others_means <- rowMeans(boot_brand[, c("Chevron.Texaco", "ExxonMobil", "Other")])
diff_mean4 = (boot_brand$Shell - boot_brand$others_means)
diff_mean4_df <- data.frame(diff_mean4)
ggplot(diff_mean4_df) + 
  geom_histogram(aes(x=diff_mean4), color= "deeppink", fill= "darkslategray")  + labs(title =  "Comparison of Gas Price Means by Brand (Shell vs. Others) (Bootstrapped)", x= "Difference in Means")
```

#### As depicted in the histogram, the peak in the difference between mean prices of Shell gas and other gas stations is approximately 0.02. While this difference may appear minor, it still suggests that Shell prices are indeed slightly higher than those of other gas brands.  ####

### Problem 2 ###

#### Part A: Filter the data set down to include only those cars where year == 2011 and trim == "63 AMG". Based on these 116 cars, compute a 95% bootstrap confidence interval for the average mileage of 2011 S-Class 63 AMGs that were hitting the used-car market when this data was collected. ####

```{r, warning=FALSE, echo=FALSE, message=FALSE}
setwd("/Users/valentinatrevino/Desktop/spring '24/SDS")
sclass_data <- read.csv("sclass.csv")

A_sclass_data <- sclass_data %>%
  filter(sclass_data$year == 2011, sclass_data$trim == "63 AMG")

mean(mileage ~ condition, data= A_sclass_data)
boot_mileage = do(10000)*mean(mileage ~ condition,
data=resample(A_sclass_data))

ggplot(boot_mileage) + 
  geom_histogram(aes(x=Used))

confint(boot_mileage, level=0.95)
```

#### Part B: Filter the data set down to include only those cars where year == 2014 and trim == "550". Based on this sample of 2889 cars, compute a 95% bootstrap confidence interval for the proportion of all 2014 S-Class 550s that were painted black. Hint: you might find this easiest if you use mutate to first define a new variable, isBlack, that is either TRUE or FALSE depending on whether the car is black. ####

```{r, warning=FALSE, echo=FALSE, message=FALSE}
B_sclass_data <- sclass_data %>%
  filter(sclass_data$year == 2014, sclass_data$trim == "550")

color_black <- B_sclass_data$color == "Black"
proportion_black <- sum(color_black) / length(B_sclass_data$color)
```

### Problem 3 ###

#### Part A. Consider the shows “Living with Ed” and “My Name is Earl.” Who makes people happier: Ed or Earl? Construct a filtered data set containing only viewer responses where Show == "Living with Ed" or Show == "My Name is Earl". Then construct a 95% confidence interval for the difference in mean viewer response to the Q1_Happy question for these two shows. Is there evidence that one show consistently produces a higher mean Q1_Happy response among viewers? ####

```{r, warning=FALSE, echo=FALSE, message=FALSE}
setwd("/Users/valentinatrevino/Desktop/spring '24/SDS")
pilot_data <- read.csv("nbc_pilotsurvey.csv")
```
